Exception Handling Bug Hazards in Android

### 2015-10-06 SPRINGER_JOURNAL/msr_journal.tex
including tables with tags and overall comments about them,organizing comments

\section{The Mining Study Design}
\label{sec:study}

@@

\section{The Mining Study Results}
\label{sec:result}

@@

\subsection{Can the root exceptions reveal bug harzards?}
After distilling the information available on the exception stack traces, we could find 
the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
This table also shows how many times the signaler of such an exception was a method defined on
the Android platform, the Android libcore, the application itself or a third-party library -
 following the classification presented in Table  \ref{tab:signalers}.

@@

\subsection{Can the exception types reveal bug hazards?}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

\subsection{Can the exception wrappings reveal bug hazards?}

@@

@@

\noindent\emph{\textbf{Q1:How often developers deal with the exception handling in Android development?}}

Many of our survey respondents said they have to handle exceptions most of the time during Android app development (64.8\%). However, the frequency on which they throw exception inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%). It shows that although we can try to avoid dealing with exceptions throwing few exceptions, it is almost always inevitable to handle the exceptions thrown by Android platform and reused libraries.

@@

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Resons} & \bfseries{\#} & \bfseries{\%} \\
\hline
\bfseries{For Positively Affecting Robustness} &   &   \\
Improves error diagnosis  &	10	& 15\% \\
Anticipated erroneous situation help writing more robust code  & 9	 & 13\% \\ 
%Anticipated erroneous situation help writing more robust code 
Exceptions only for unrecoverable behavior	& 7 &	10\% \\
Checked exceptions force handling  &	6	& 9\% \\
Useful to gracefully crash 	& 4 &	6\% \\

@@

@@

///// PREVENT CRASHES /////
Finally developers were asked how they prevent their apps from crashing. Most of the respondents answered that they prevent apps from crashing by handling exceptions/errors (27\%). They mentioned about some strategies such as: handling exceptions in all entry points/all over; using catch-all clauses; and adding try catch blocks around risky methods. Different from conventional Java programs, an Android app is composed by several entry points, each window (i.e. Activity) is a potential entry point on which exceptions can scape leading to an app crash. Hence the importance of handling exceptions in all entry points, and using general cash clauses which cannot sometimes prevent crash but allows the developer to present a detailed error message to the user before crashing. Many respondents also said that they used testing approaches to prevent crashes -  few of them mentioned crowd testing (3\%) an emerging approach in testing apps. Moreover, null checks were also mentioned by 15\% of the developers as a way to prevent crashes. The adoption of coding styles were also pointed out by 14\% of respondents as a way to prevent craches. Finally, the crash report tools (e.g. ACRA [ref]) were reported as a way to prevent crashes. Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented. Table ~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.	////

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
\hline
exception/error handling & 19 &	27\% \\
testing &	14 &	20\% \\
null check(s) &	11 &	15\% \\
- null check annotations (@Null and @NonNull)	& 3 & 	4\% \\
coding style(s)	& 10 &	14\% \\
crash report / crash report tools &	8	& 11\% \\
\hline
\end{tabular}
\caption{Top ways of preventing crashes - 71 non-empty respondents. }
\label{tab:topprevent}
\end{table}

/////// JAVA BEST PRACTICES ////// 
As mentioned before several general guidelines have been proposed on how to
use Java exceptions. Section XXXXX compiled a list some exception handling best practices that guided our mining study. In this survey also we asked the Android developers whether they knew about any Java exception handling best practice. Although most of them said that knew and adopted EH practices 68\%, from this set the most mentioned best practices are presented on Table XXXX - considering that one developer mentioned more than one practice. Some of the practices mentioned, however, are not well known Java EH best practices but seam to be used by some Android developers to better deal with exceptions in Android app context: the crash fast approach [ref], the use of crash report tools, and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps./////

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
\hline
Use specific handlers / dont catch generic exceptions	& 9 & 23\% \\
Don't swallow Exceptions	 & 7 & 18\% \\
Dont throw Runtime / Favor Checked exceptions & 4 &	10\% \\
Do not use exception for normal  flow control. &	4 &	10\% \\
Free Resources in finally-blocks	& 4	& 10\% \\
crash fast	 & 3 &	8\% \\
crash report tools	&  2 &	5\% \\
Don't catch Errors 	&  2 &	5\% \\
\hline
\end{tabular}
\caption{Top mentioned Java EH best practices -  40 non-empty answers. }
\label{tab:javapractices}
\end{table}

/////// ANDROID EH BEST PRACTICES /////
Developers who informed that adopted EH best practices were also asked whether they knew about any EH best practice specific to Android context. Most of them 43\% mentioned that they applied the same best practices that they used in Java programs. Only some of them mentioned that adopted specific best practices such as: (i) the use of crash report tools (21\%)  - to notify the developer about the uncaught exceptions that hapen in the application code; and (ii) the use of a global exception handler (14\%) - the UncaughtExceptionHandler advised on Android tutorials [ref].	 /////

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
\hline
same as java	& 6 &	43\% \\
crash report tools	 & 3 &	21\% \\
add a global exception handler (UncaughtExceptionHandler)	 & 2 & 14\% \\
use checked exceptions	& 1 &	7\% \\
use appropriate exception messages	& 1 &	7\% \\
\hline
\end{tabular}
\caption{Top mentioned Android-specific EH best practices -  14 non-empty responses. }
\label{tab:androidpractices}
\end{table}

///// HANDLING RUNTIME AND CHECKED /////
Developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods. Then de developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception. This question was motivated by the fact that during the mining study we could identify that many checked exception were just wrapped in a runtime exception and rethrown, as a way to by pass any kind of handling. When asked about of how to deal with a checked exception signaled by a method (invoked in the context of onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation. Also most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such method by a try-catch block (xxxxx\%). However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases. One of them mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled. Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user. On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception. Some developers also mentioned to use a Toast to support this task - a toast is an Android component that provides simple feedback about an operation in a small popup. (http://developer.android.com/guide/topics/ui/notifiers/toasts.html). Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app. Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so. //////

@@

@@

>>>>

\section{The Repository Mining Study}
\label{sec:study}

@@

\subsection{The Mining Study Results}
\label{sec:result}

@@

\subsubsection{Can the root exceptions reveal bug harzards?}
After distilling the information available on the exception stack traces, we could find 
the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
This table also shows how many times the signaler of such an exception was a method defined on
the Android platform, the Android libcore, the application itself or a third-party library -
 following the classification presented in Table  \ref{tab:signalers}.

@@

\subsubsection{Can the exception types reveal bug hazards?}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

\subsubsection{Can the exception wrappings reveal bug hazards?}

@@

\bigskip
\noindent\emph{\textbf{Q1:How developers deal with the exception handling in Android development?}}
\bigskip

This first research question explores the ways in which developers deal with the exception handling code, either throwing or handling exceptions,  while developing an Android app.  Developers were asked about the frequency they develop exception handling code and whether or not they adopt best practices while developing.

%\noindent\emph{Frequency}

Developers were firstly asked how often they dealt with the exception handling code. Many of our survey respondents recognized they have to handle exceptions most of the time during Android app development (64.8\%). However, the frequency on which they throw exception inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%). It shows that although we can try to avoid dealing with exceptions throwing few exceptions, it is almost always inevitable to handle the exceptions thrown by Android platform and reused libraries.

%\emph{EH Best Practices}
%/////// JAVA BEST PRACTICES ////// 
As mentioned before several general guidelines have been proposed on how to
use Java exceptions. Section XXXXX compiled a list some exception handling (EH) best practices that guided our mining study. In this survey the developers were also asked if they knew about and used Java EH best practices. Although most of them said that knew and adopted Java EH practices 68\%, from this set the most mentioned best practices are presented on Table XXXX - considering that one developer mentioned more than one practice. Some of the practices mentioned, however, are not well known Java EH best practices but seam to be used by some Android developers to better deal with exceptions in Android app context: the crash fast approach [ref], the use of crash report tools, and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps.

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Java EH Best Practices} & \bfseries{\#} & \bfseries{\%} \\
\hline
Use specific handlers / dont catch generic exceptions	& 9 & 23\% \\
Don't swallow Exceptions	 & 7 & 18\% \\
Dont throw Runtime / Favor Checked exceptions & 4 &	10\% \\
Do not use exception for normal  flow control. &	4 &	10\% \\
Free Resources in finally-blocks	& 4	& 10\% \\
crash fast	 & 3 &	8\% \\
crash report tools	&  2 &	5\% \\
Don't catch Errors 	&  2 &	5\% \\
\hline
\end{tabular}
\caption{Top mentioned Java EH best practices -  40 non-empty answers. }
\label{tab:javapractices}
\end{table}

%/////// ANDROID EH BEST PRACTICES /////
The developers who informed that adopted EH best practices were also asked whether they knew about any EH best practice specific to Android context. Most of them 43\% mentioned that they applied the same best practices that they used in Java programs. Only some of them mentioned that adopted specific best practices such as: (i) the use of crash report tools (21\%)  - to notify the developer about the uncaught exceptions that hapen in the application code; and (ii) the use of a global exception handler (14\%) - the UncaughtExceptionHandler advised on Android tutorials [ref].

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Android EH Best Practices } & \bfseries{\#} & \bfseries{\%} \\
\hline
Same as java	& 6 &	43\% \\
Crash report tools	 & 3 &	21\% \\
Add a global exception handler (UncaughtExceptionHandler)	 & 2 & 14\% \\
Use checked exceptions	& 1 &	7\% \\
Use appropriate exception messages	& 1 &	7\% \\
\hline
\end{tabular}
\caption{Top mentioned Android-specific EH best practices -  14 non-empty responses. }
\label{tab:androidpractices}
\end{table}

@@

\noindent\emph{\textbf{Q2: How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}}

@@

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Reasons} & \bfseries{\#} & \bfseries{\%} \\
\hline
\bfseries{For Positively Affecting Robustness} &   &   \\
Improves error diagnosis  &	10	& 15\% \\
Anticipated erroneous situation help writing more robust code  & 9	 & 13\% \\ 
%Anticipated erroneous situation help writing more robust code 
Exceptions only for unrecoverable behavior	& 7 &	10\% \\
Checked exceptions force handling  &	6	& 9\% \\
Useful to gracefully crash 	& 4 &	6\% \\

@@

%\bigskip

Finally developers were asked how they prevent their apps from crashing. Most of the respondents answered that they prevent apps from crashing by handling exceptions/errors (27\%). They mentioned about some strategies such as: handling exceptions in all entry points/all over; using catch-all clauses; and adding try catch blocks around risky methods. Different from conventional Java programs, an Android app is composed by several entry points, each window (i.e. Activity) is a potential entry point on which exceptions can scape leading to an app crash. Hence the importance of handling exceptions in all entry points, and using general cash clauses which cannot sometimes prevent crash but allows the developer to present a detailed error message to the user before crashing. Many respondents also said that they used testing approaches to prevent crashes -  few of them mentioned crowd testing (3\%) an emerging approach in testing apps. Moreover, null checks were also mentioned by 15\% of the developers as a way to prevent crashes. The adoption of coding styles were also pointed out by 14\% of respondents as a way to prevent craches. Finally, the crash report tools (e.g. ACRA [ref]) were reported as a way to prevent crashes. Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented. Table ~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top Ways of Crash Prevention} & \bfseries{\#} & \bfseries{\%} \\
\hline
handling exception/error  & 19 &	27\% \\
testing &	14 &	20\% \\
null check(s) &	11 &	15\% \\
- null check annotations (@Null and @NonNull)	& 3 & 	4\% \\
coding style(s)	& 10 &	14\% \\
crash report / crash report tools &	8	& 11\% \\
\hline
\end{tabular}
\caption{Top ways of preventing crashes - 71 non-empty respondents. }
\label{tab:topprevent}
\end{table}

//// research

@@

\noindent\emph{\textbf{Q3.1: How NullPointerException impact the development of robust Android apps?}}

@@

\noindent\emph{\textbf{Q3.2: How Cross-Type Wrappings impact the development of robust Android apps?}}

Some of the best practices are related to the different ways of handling checked and unchecked exceptions. When the designer of an API specifies that a method throws a checked exception, it is telling
the caller of the API that such exception should be handled.

///// HANDLING RUNTIME AND CHECKED /////
Developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods. Then de developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.

This question was motivated by the fact that during the mining study we could identify that many checked exception were just wrapped in a runtime exception and rethrown, as a way to by pass any kind of handling. When asked about of how to deal with a checked exception signaled by a method (invoked in the context of onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation. Also most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such method by a try-catch block (xxxxx\%). However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases. One of them mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled. Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user. On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception. Some developers also mentioned to use a Toast to support this task - a toast is an Android component that provides simple feedback about an operation in a small popup. (http://developer.android.com/guide/topics/ui/notifiers/toasts.html). Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app. Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so. //////

@@

\bigskip
\noindent\emph{\textbf{RQ 3.3 Are developers aware the hidden dangers of non-declared checked exceptions?}}
\bigskip
